Think deep for reef resilience

### 2025-08-16 paper/paper_draft.qmd
Add note for all authors!
Correct timeseries clustering timespan to 2025-2075

The time series clustering was applied over the time frame between 2030 and 2060. The
selected time frame prevents the classification from being overly influenced by any steep
declines in the initial years of simulations, and avoids the post-2060 time period where
reefs are projected to be under severe stress and in decline [@bozec2025]. Spatially, the
analyses were conducted for each GBR Management area and at the scale of each reefal
bioregion (defined in GBRMPA [-@GBRMPA2024]). Analyses were also performed at the
whole-of-GBR scale but were found to lead to identical results **\[TI: Ben, compared to
what, the bioregion or the Management area results?\]** and so are not shown here. To
compare time series clustering results across GCMs, Management area scale clustering results
are used to preserve coarse scale latitudinal differences between regions, while maintaining
ease of comparison and visualisation. These comparisons include quantification of
differences between low and high cover clusters across management regions under each GCM
using the CID between cluster median time series. To assess how different reef
characteristics influence time series clustering at a finer spatial scale, bioregion results
are used to capture patterns occurring within individual clusters of reefs. Results of all
analyses at all clustering scales can be found at Grier and Iwanaga [-@grier2025] for those
interested.

@@

@@

The process creates a distance matrix of n \* n dimensions (where n is the number of time
series in the group) that is clustered using a k-medoids clustering algorithm [cf.
@steinmann2020]. The three reef clusters are assigned a qualitative categories from Low,
Medium and High.

>>>>

The time series clustering was applied over the time frame between 2025 and 2075. Spatially,
the analyses were conducted for each GBR management area and at the scale of each reefal
bioregion (defined in GBRMPA [-@GBRMPA2024]). Analyses were also performed at the
whole-of-GBR scale but were found to lead to identical results **\[TI: Ben, compared to
what, the bioregion or the Management area results?\]** and so are not shown here. To
compare time series clustering results across GCMs, Management area scale clustering results
are used to preserve coarse scale latitudinal differences between regions, while maintaining
ease of comparison and visualisation. These comparisons include quantification of
differences between low and high cover clusters across management regions under each GCM
using the CID between cluster median time series. To assess how different reef
characteristics influence time series clustering at a finer spatial scale, bioregion results
are used to capture patterns occurring within individual clusters of reefs. Results of all
analyses at all clustering scales can be found at Grier and Iwanaga [-@grier2025] for those
interested.

@@

**\[TI note: Important to speak to interesting depth range at edge of where coral planting
might feasibly/viably occur with current approaches\]**

@@

The process creates a distance matrix of n \* n dimensions (where n is the number of time
series in the group) that is clustered using a k-medoids clustering algorithm [cf.
@steinmann2020]. The three reef clusters are labelled with qualitative-ordered categories
from Low, Medium and High based on their median cover relative to each other cluster over
the 2030-2060 period.The selected time frame prevents the classification from being overly
influenced by any steep declines in the initial years of simulations, and avoids the
post-2060 time period where reefs are projected to be under severe stress and in a declined
state [@bozec2025].
